---
title: "Machine Learning_hw1_jf3354"
author: Judy Fordjuoh
date: January 21, 2022
output: html_document
---

```{r, echo=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(haven)
library(knitr)
library(broom)
library(ggplot2)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .9,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## QUESTION 1
```{r}
#Task: Construct a table providing summaries of the quantitative features of the dataset. Summaries should include the mean, median, minimum value, and maximum value. If you are unable to construct a formatted table within R, you can print raw output, but then comment the output to identify the answer that was requested.

bcdata = read_csv("bcdata_Assignment1.csv") %>%
    janitor::clean_names()
summary(bcdata)
  
```
For the age category, the minimum age in the dataset is 24 years, the median is 56 years, the average age (mean) is 57.3 years, and the maximum age in this data set is 89 years. In the BMI category,the lowest BMI in the dataset is 18.37 kg/m2, the median BMI is 27.66 kg/m2, the average BMI is 27.58 kg/m2, and the highest BMI in this data set is 38.58 kg/m2. As for glucose, the lowest glucose level is 60.0 mg/dL, the median is 92.00 mg/dL, the average glucose level is 97.79 mg/dL, and the maximum glucose level in this dataset is 201.00 mg/dL. The lowest insulin level is 2.432 μU/mL, the median is 5.925 μU/mL, the average level is 10.012 μU/mL, and the maximum insulin level is 58.460 μU/mL. The lowest result from the homeostatic model assessment was 0.4674, the median was 1.3809, the mean was 2.6950, and the maxium result in this dataset was 25.0503. The lowest level of leptin was 4.311 ng/mL, the median was 20.271 ng/mL, the average level was 26.615 ng/mL, and the maximum level of leptin in this dataset was 90.280 ng/mL. The lowest level of adiponectin was 1.656 μg/mL, the median was 8.353 μg/mL, the average level was 10.181 μg/mL, and the maximum level of adiponectin in this dataset was 38.040 μg/mL. The lowest level of resistin was 3.210 ng/mL, the median was 10.828 ng/mL, the average level was 14.726 ng/mL, and the maximum level of resistin in this dataset was 82.100 ng/mL. The lowest chemokine level (MCP-1) was 45.84 pg/dL, the median was 471.32 pg/dL, the average level was 534.65 pg/dL, and the maximum chemokine level in this dataset was 1698.44 pg/dL. Since the classification labels is only two levels (1= controls and 2= breast cnacer patients), the minimium = 1, the maximum =2, mean = 1.552, and the median = 2. Since the classification labels is mainly used to group the patients, this descriptive explaniation of this variable may not be as useful.


## QUESTION 2
```{r}
#Task: Recode BMI into the WHO-defined categories below: Severely underweight - BMI less than 16.5kg/m^2, Underweight - BMI under 18.5 kg/m^2, Normal weight - BMI greater than or equal to 18.5 to 24.9 kg/m^2, Overweight – BMI greater than or equal to 25 to 29.9 kg/m^2, Obesity class I – BMI 30 to 34.9 kg/m^2, Obesity class II – BMI 35 to 39.9 kg/m^2, Obesity class III – BMI greater than or equal to 40 kg/m^2. 

#Note to self: there is no severely underweight in this data set
 
#Note to self: regrouping the BMI, referred to DS midterm

bcdata = bcdata %>%  
  mutate(bmi_category = cut(bmi, breaks = c(0, 16.5, 18.5, 24.9, 29.9, 34.9, 39.9, 50), labels = c("Severely Underweight","Under Weight", "Normal Weight","Overweight","Obesity Class  I","Obesity Class II", "Obesity Class III"))) %>%  
  select(bmi,bmi_category, everything()) 

```
After categorizing the individuals to the respective BMI categories, we see that there is no study participants in the Severely Underweight and the Obesity Class III BMI categories. 


## QUESTION 3
```{r}
#Task: Create a bar chart showing the proportion of breast cancer cases and controls within each BMI category

#Note to self: referred to DS midterm

dataclean2 = bcdata %>% 
  mutate(classification = as.character(classification)) %>%
  mutate(classification = recode(classification,
                       "1" = "Healthy Controls",
                       "2" = "Breast Cancer Patients")) %>% 
  group_by(bmi_category, classification) %>%
  summarize(total = n()) %>%
  mutate(proportion = total/sum(total)) %>% #the proportion in 
  mutate(percentage = 100*proportion)
  
bcdata_graphs <- dataclean2 %>%
  ggplot(aes(x = classification, y = percentage, fill = classification)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d() +
  facet_grid(~bmi_category) +
  labs(
    title = "Proportion of Breast Cancer Cases and 
    Controls Within Each BMI Category (in percentage)",
    x = "Patient Type",
    y = "Percentage of Study Participants"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "bottom")

bcdata_graphs


```

We see that in the underweight category, there is only one breast cancer patient (100%) and no healthy controls. For the Normal Weight, Overweight, and Obesity Class I categories, the breast cancer patients surpass the healthy controls. In the Obesity Class II group, there are more healthy controls in comparison to the breast cancer patients. There were no participants in the severely underweight and obesity class III group, which was reflected in the graphs.   


## QUESTION 4
```{r}
# Construct a logistic regression model using breast cancer classification as the outcome and glucose, HOMA-IR, leptin, BMI (continuous) and age as the independent variables. Fill in the beta estimate and 95% confidence interval associated with a 1-unit change in HOMA-IR

bc = bcdata %>% 
mutate(classification = as.factor(classification)) 

logistic = glm(classification ~ glucose + homa + leptin + bmi + age, data = bc, family = binomial()) 
summary(logistic)

logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3) #The table below summaries the coefficients from the model fit; because logistic model estimates are log odds ratios, we include a step to compute odds ratios as well.

confint.default(logistic, level = 0.95, type = Wald)

#calculating the 95% CI
  #tidy(conf.int = T)
```
HOMA Beta: 0.273882
Interpretation: For every one unit change in the homeostatic model assessment, the log odds of being a breast cancer patient increases by 0.274. 

Interpretation 2: The log odds of being a breast cancer patient with a high HOMA level is 0.274 times the log odds of being a breast cancer patient with a low HOMA level. 

HOMA CI: (-0.063, 0.611) 
We are 95% confident that the log odds of being a breast cancer patient comparing those with a high HOMA level to those with a low HOMA level, lies between -0.063 and 0.611.


## QUESTION 5
```{r}
#Construct a linear regression model using insulin as the outcome and BMI (continuous), age, and glucose as the independent variables. Fill in the beta estimate and 95% confidence interval associated with a 1-unit change in age.

linear = lm(insulin ~ bmi + age + glucose, data = bcdata)

summary(linear) #this would tell you the call, the residuals, the coefficients, SE, Rsquared and F stat

broom::tidy(linear) %>%
  select(-std.error, -statistic) %>%
  knitr::kable(digits = 3)

confint.default(linear, level = 0.95, type = Wald)
```
Age Beta:-0.054
For every one unit change in the age, the amount of insulin being produced decreases by 0.054. 

Age CI: (-0.156,0.048)
We are 95% confident that the   of being a breast cancer patient comparing those with a high HOMA level to those with a low HOMA level, lies between -0.063 and 0.611.